---
title: "2022-03-13: Myelin phagocytosis"
output: html_notebook
---

```{r include=FALSE}
library(aomisc) # remotes::install_github("onofriAndreaPG/aomisc")
library(tidyverse)
library(magrittr)
library(readxl)
library(broom)
library(patchwork)
library(lmerTest)
library(DHARMa)
theme_set(theme_minimal())
```

## Data

Macrophages derived from Control (CT) and PD individuals were exposed to myelin prepared from CT individuals.

```{r}
df = read_xlsx("ctrl_myelin-macrophages.xlsx")

df

df %<>%
  pivot_longer(-TIME, values_to = "INT", names_to = "ID") %>%
  separate(ID, sep = 2, into = "STATUS", remove = FALSE) %>%
  select(ID, STATUS, TIME, INT) %>%
  arrange(ID, TIME)

df
```

```{r}
ggplot(data = df, mapping = aes(x = TIME, y = INT, color = STATUS)) + geom_smooth(method = "loess", formula = y ~ x)
```
```{r}
ggplot(data = df, mapping = aes(x = TIME, y = INT, group = ID, color = STATUS)) +
  geom_point(size = 1) +
  geom_line(alpha = 0.25)
```

## Linear regression forcing intercept to zero

```{r}
ggplot(data = df, mapping = aes(x = TIME, y = INT, group = ID, color = STATUS)) +
  geom_line(stat = "smooth", method = "lm", formula = y ~ 0 + x, se = FALSE, alpha = 0.5)
```

```{r}
m1.fit = lmer(INT ~ 0 + TIME*STATUS + (0 + TIME | ID), data=df)

summary(m1.fit)
```

```{r}
m1.res = simulateResiduals(fittedModel = m1.fit)
plot(m1.res)
```
## Linear regression forcing intercept to zero and up to 10 hours

```{r}
m2.fit = lmer(INT ~ 0 + TIME*STATUS + (0 + TIME | ID), data=df %>% filter(TIME <= 10))

summary(m2.fit)
```

```{r}
m2.res = simulateResiduals(fittedModel = m2.fit)
plot(m2.res)
```

## Linear regression not forcing intercept to zero

```{r}
ggplot(data = df, mapping = aes(x = TIME, y = INT, group = ID, color = STATUS)) +
  geom_line(stat = "smooth", method = "lm", formula = y ~ x, se = FALSE, alpha = 0.5)
```


```{r}
m3.fit = lmer(INT ~ TIME*STATUS + (TIME | ID), data=df)

summary(m3.fit)
```

```{r}
m3.res = simulateResiduals(fittedModel = m3.fit)
plot(m3.res)
```
## Linear regression not forcing intercept to zero and up to 10 hours

```{r}
m4.fit = lmer(INT ~ TIME*STATUS + (TIME | ID), data=df %>% filter(TIME <= 10))

summary(m4.fit)
```

```{r}
m4.res = simulateResiduals(fittedModel = m4.fit)
plot(m4.res)
```

## Two-factor repeated-measures ANOVA

```{r}
ggplot(data = df, mapping = aes(x = factor(TIME), y = INT, color = STATUS)) +
  geom_boxplot()
```
```{r}
summary(aov(INT ~ STATUS*factor(TIME) + Error(ID), data = df))
```


```{r}
anova(lmer(INT ~ STATUS*factor(TIME) + (1 | ID), data=df))
```

```{r}
summary(lmer(INT ~ STATUS*factor(TIME) + (1 | ID), data=df))
```

## Model curve using aomisc package

https://www.statforbiology.com/2020/stat_nls_usefulfunctions/

```{r}
m = df %>%
  nest(data = -c(ID, STATUS)) %>%
  mutate(
    model = map(.x = data, ~ drm(.x$INT ~ .x$TIME, fct = DRC.negExp())),
    tidied = map(model, tidy),
    predicted = map(model, predict)
  )

t = m %>% 
  unnest(tidied) %>%
  pivot_wider(id_cols = c(ID, STATUS), names_from = term, values_from = estimate) %>%
  rename(PLATEAU = a, RATE = c)

t
```


```{r}
t.test(PLATEAU ~ STATUS, data = t)
wilcox.test(PLATEAU ~ STATUS, data = t)
```

```{r}
t.test(RATE ~ STATUS, data = t)
wilcox.test(RATE ~ STATUS, data = t)
```

```{r}
df2 = df %>% left_join(m %>%
    mutate(TIME = list(0:20)) %>%
    select(ID, TIME, predicted) %>%
    unnest(cols = c(TIME, predicted)),
  by = c("ID", "TIME"))
```

```{r}
(ggplot(data = df2, mapping = aes(x = TIME, y = INT, group = ID, color = ID)) +
  geom_point(size = 1) +
  geom_line(alpha = 0.25) +
  ylim(0, 300000000) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none") +
ggplot(data = df2, mapping = aes(x = TIME, y = INT, group = ID, color = ID)) +
  geom_point(size = 1) +
  geom_line(mapping = aes(y = predicted), alpha = 0.25) +
  ylim(0, 300000000) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "none")) / 
(ggplot(data = df2, mapping = aes(x = TIME, y = INT, color = STATUS)) +
  geom_line(mapping = aes(group = ID), stat="smooth", method = "lm", formula= y ~ 0 + x, se = FALSE, alpha = 0.25) +
  geom_smooth(method = "loess", formula="y ~ x", linetype = "dashed", se = FALSE) +
  ylim(0, 300000000) +
  theme(legend.position = "none") + 
 ggplot(data = df2, mapping = aes(x = TIME, y = INT, color = STATUS)) +
  geom_line(mapping = aes(y = predicted, group = ID), alpha = 0.25) +
  geom_smooth(method = "loess", formula = y ~ x, linetype = "dashed", se = FALSE) +
  ylim(0, 300000000) +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none"))
```

---

```{r}
library(saemix)

saemix.data <- saemixData(name.data       = df,
                          name.group      = "ID",
                          name.predictors = "TIME",
                          name.response   = "INT")
```

```{r}
model.fn <- function(psi, id, x) { 
  t = x[, 1] 
  plateau = psi[id, 1]
  rate = psi[id, 2]
  pred = plateau * (1 - exp(-rate * t))
  return(pred)
}

saemix.model <- saemixModel(model = model.fn, 
                            psi0  = c(plateau = 10000000, rate = 0.1))
```

```{r}
saemix.fit1 = saemix(saemix.model, saemix.data)
```

See also:

- https://stats.oarc.ucla.edu/r/seminars/repeated-measures-analysis-with-r/
- http://sia.webpopix.org/nlme.html
- https://saemixr.github.io/
- https://doi.org/10.1213%2FANE.0000000000003511
- https://cran.r-project.org/web/packages/growthrates/vignettes/Introduction.html
- http://dpmartin42.github.io/posts/Piecewise-growth
- https://doi.org/10.1016/j.buildenv.2021.108349
- https://doi.org/10.3758/s13423-017-1281-0